Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches

### 2017-11-09 paper.tex
noice
okay

Chick et al. showed that by significantly relaxing the precursor mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in identified peptide mass and precursor mass.
A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for open searches.
Search times for $\pm 500$ Da precursor tolerance went from six years to six seconds.
The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, with interpretation of the mass difference left as an exercise for the user.

@@

We identified peaks in the original G-PTM-D approach with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of sulfonation and phosphorylation, and a problematic peak at mass of Leucine, with a high FDR (see Table above).
INSERT A GREAT PLOT HERE We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
MSFragger also doesn’t automatically calibrate spectra prior to searching.
Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.

>>>>

Chick et al. showed that by significantly relaxing the precursor mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in identified peptide mass and precursor mass.
A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for open searches.
$\pm 500$ Da precursor tolerance searches saw a 100-fold improvement in speed over most existing proteome database search tools.
The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, with interpretation of the mass difference left as an exercise for the user.

@@

We identified peaks in the original G-PTM-D approach with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of sulfonation and phosphorylation, and a problematic peak at mass of Leucine, with a high FDR (see Table above).
INSERT A GREAT PLOT HERE We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
MSFragger also doesn't automatically calibrate spectra prior to searching.
Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
